---
output: github_document
---

# Differentially Private Kruskal Wallis Application
#### *USA Findex Data -- Age and Income Quantile*

This file analyzes the power of our DP Kruskal-Wallis test on real-world data.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(latex2exp)
library(rmutil)
library(readxl)
```

This file serves to carry out the real-world application of our differentially private Kruskal-Wallis test.

```{r import}
findex <- read_excel("kw/data/usa_findex.xls")
```

`findex` includes 1005 observations of 81 variables about American citizens' finances. Findex tables are useful for quantifying financial inclusion in an economy. In this analysis, we'll focus on the `inc_q` and `age` variables, which represent the income quantile and age of the individual, respectively.

```{r}
findex <- findex %>%
  select(inc_q, age)

findex$inc_q <- case_when(findex$inc_q == "Poorest 20%" ~ 1,
                          findex$inc_q == "Second 20%" ~ 2,
                          findex$inc_q == "Middle 20%" ~ 3,
                          findex$inc_q == "Fourth 20%" ~ 4,
                          findex$inc_q == "Richest 20%" ~ 5)
findex$age <- as.numeric(findex$age)

findex %>%
  group_by(inc_q) %>%
  summarize(age = mean(age)) %>%
  ggplot() + 
    geom_col(aes(x = inc_q, y = age)) + 
    theme_minimal() + 
    labs(x = "Income Quantile", y = "Age")
```

We can see that, as income quantile increases, the mean age in each group increases. We'll use both the public Kruskal Wallis and our differentially-private version to see if this difference is statistically significant.

First, we'll load the necessary functions. See the `.R` files for function documentation.

```{r source}
source('kw/fxns/kw_new.R')
source('kw/fxns/kw.R')
source('kw/fxns/gen_data.R')
source('kw/fxns/calc_stat.R')
```

We want to select many random samples of size `sizes` from this data, and, for each sample, run it through both the public and private Kruskal Wallis. The `pwr_plot` function carries out this whole process, with inputs as follows.

* `sizes`: A list of numeric vectors of length x, indicating the sample sizes to carry out the test with. Each element of the list must be the same length.
* `epsilons`: A list of numerics (where NA indicates a public test)
* `kw_version`: A character vector: either "sq" (referring to the regular H statistic), "abs" (referring to the unsquared absolute value H statistic), or both, so that noise with the appropriate sensitivity can be generated.
* arguments beginning with `mean` and `sd` are vectors of length x, indicating the parameters to generate data for the alternate and null distribution for.
* `reps`: A numeric indicating the number of repititions to carry out the simulations with.
* `alpha`: A numeric indicating the significance level to carry the test out with.

```{r}
pwr_plot <- function(sizes, epsilons, kw_version, mean_null, sd_null, reps, alpha) {
   
 # We first create a list of lists, (can be thought of as a dataframe where
 # each row is a unique combination of sizes, epsilons, and kw_version). We
 # do this step instead of using cross() for clarity of order.
  arg_list <- list(n =          rep(sizes,
                                    each = length(epsilons)*length(kw_version)), 
                   epsilon =    rep(epsilons,
                                    times = length(sizes)*length(kw_version)),
                   kw_version = rep(kw_version,
                                    times = length(sizes),
                                    each = length(epsilons)))
  
# We need to do the same for the simulation of the reference distribution,
# where we assume equal group sizes
  null_sizes_1d <- lapply(sizes, mean) %>% lapply(., ceiling)
  null_sizes    <- lapply(1:length(null_sizes_1d), 
                          function(x) {rep(null_sizes_1d[[x]], length(sizes[[1]]))})

  arg_list_null <- list(n =     rep(null_sizes,
                                    each = length(epsilons)*length(kw_version)), 
                   epsilon =    rep(epsilons,
                                    times = length(sizes)*length(kw_version)),
                   kw_version = rep(kw_version,
                                    times = length(sizes),
                                    each = length(epsilons)))
    
    
# Simulate null distributions with varying parameters and then calculate test 
# statistics from the null distributions. the output, null_stats is a 
# dataframe, where each column is a distinct combination of epsilon, sample 
# size, and kw_version.
  
  null_data <- pmap(.l = arg_list_null,
                 .f = gen_data,
                 mean = mean_null,
                 sd = sd_null, 
                 reps = reps)
  null_stats <- map_dfc(.x = null_data , .f = calc_stat)

# Calculate critical values for each combination of epsilon, sample 
# size, and kw_version.
  
  crit_vals  <-   map(null_stats, quantile, probs = 1 - alpha) %>% 
                  unlist(use.names = FALSE)
  
# This is where the procedure will diverge from that in the simulation case. Instead of
# simulated data, we would like to use data from `findex`. `calc_stat` takes in a dataframe
# with formatting like that outputted by `gen_data`. We're simply "switching out" the 
# x and group columns with new values from findex_samples
  findex_samples <- list()
  alt_data <- list()
  
  for (i in 1:length(null_data)) {
    findex_samples[[i]] <- list(sample_n(findex, 
                                         nrow(null_data[[i]][[1]]),
                                         replace = TRUE), 
                                null_data[[i]][[2]])
    
    alt_data[[i]] <- list(cbind(null_data[[i]][[1]], findex_samples[[i]][[1]]),
                          null_data[[i]][[2]])
    
    alt_data[[i]][[1]] <- alt_data[[i]][[1]] %>%
      select(-x, -group) %>%
      rename(x = age, group = inc_q)
  }

# We now want to calculate test statistics from all of the dataframes in alt_data.  
  alt_stats <- map_dfc(.x = alt_data, .f = calc_stat)
  
# this is a vector of powers at several parameterizations
   
  power_vector <- map2(.x = crit_vals, .y = alt_stats, .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)

  
# make a list of the total sample size for each element of sizes 
  N_list <- rep(lapply(sizes, sum), 
                each = length(epsilons)*length(kw_version))  
  
# put together a dataframe of powers resulting from given parameters
  plot_df <- cbind(N = unlist(N_list),
                   bind_rows(arg_list),
                   power_vector) %>%
             as_data_frame()
  
# check if there are NA's in epsilon; if so, convert to "Public" for plotting
  plot_df$epsilon[is.na(plot_df$epsilon)] <- "Public"

  plot_df
}
```

We now make use of the `pwr_plot` function to put together a dataframe, plot it, and save it.

```{r usage}
# make a vector of sample sizes
sizes_1d <- seq(from = 1.1, to = 2.15, by = .05)^10 %>% round() %>% as.list()
sizes <- lapply(1:length(sizes_1d), function(x) {rep(sizes_1d[[x]], 5)})

# simulate the plot data
kw_application <- pwr_plot(sizes = sizes,
         epsilons = list(.1, 1, NA),
         kw_version = c("sq", "abs"),
         mean_null = c(0, 0, 0, 0, 0),
         sd_null = c(1, 1, 1, 1, 1),
         reps = 1e4,
         alpha = .05)

# save the plot data
save(kw_application, file = "kw/data/kw_application.Rda")

# create a greyscale-friendly, qualitative color palette
colors <- c("#000099", "#dd02e0", "#ece900", "#05e200")

# plot the data
kw_application_plot <- ggplot(kw_application) + 
    geom_line(aes(x = unlist(N), 
                  y = unlist(power_vector), 
                  col = as.factor(unlist(epsilon)),
                  linetype = factor(kw_version)),
              size = 1.3) +
    labs(x = "Total Sample Size n", y = "Power") +
    scale_color_manual(name = TeX("$\\epsilon$"),
                       values = colors) +
    scale_linetype_discrete(name = "Test Type",
                            labels = c(unname(TeX("KW_{abs}")), "KW")) +
    scale_x_log10() +
    theme_minimal(base_family = "Times",
                  base_size = 24)  + 
    theme(axis.line = element_line(colour = "black"))

kw_application_plot

# save the plot
ggsave(filename = "kw_application.png", 
       plot = kw_application_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```

