---
output: github_document
---
# Kruskal-Wallis Power Analysis
#### *Uniformity of p-values*

This file analyzes the error rate of our DP Kruskal-Wallis test using quantile-quantile plots.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(rmutil)
library(latex2exp)
```

Some of the functions used in this document are used, unedited, in several other documents throughout the power analysis. We load them now---documentation for each can be found in the relevant files in the `fxns` folder.

```{r}
source('kw/fxns/kw.R')
source('kw/fxns/kw_new.R')
source('kw/fxns/calc_stat.R')
source('kw/fxns/gen_data.R')
```

`qq_plot` is a wrapper function that vectorizes the functions defined above to make a dataframe for our qqplots.

```{r qq_plot}
qq_plot <- function(sizes, alphas, epsilon, kw_version, mean, sd, reps) {
   
# make a new sizes list to generate the null distribution that assumes
# equal group sizes
  null_sizes_1d <- lapply(sizes, mean) %>% lapply(., ceiling)
  null_sizes    <- lapply(1:length(null_sizes_1d), 
                          function(x) {rep(null_sizes_1d[[x]], 3)})
  
# simulate distributions of test statistics at each N 
  null_stats <- map(.x = null_sizes,
                    .f = gen_data,
                    mean = mean,
                    sd = sd,
                    reps = reps,
                    epsilon = 1,
                    kw_version = kw_version) %>%
                map_dfc(.x = ., .f = calc_stat)
  
# calculate critical values for each alpha and N  
  crit_vals  <-   map(null_stats, quantile, probs = c(1 - alphas)) %>% 
                  unlist(use.names = FALSE)
  
# compute "alternate" statistics (with the same parameters),
# except the true sample sizes are used
  alt_stats <- map(.x = sizes,
                    .f = gen_data,
                    mean = mean,
                    sd = sd,
                    reps = reps,
                    epsilon = 1,
                    kw_version = kw_version) %>%
                map_dfc(.x = ., .f = calc_stat)
  
# compute a vector of type I error rates
  type1 <- map2(.x = crit_vals, 
                       .y = rep(alt_stats, 
                                each = length(alphas)), 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)

# make a list of the total sample size for each element of sizes 
  N_list <- rep(lapply(sizes, sum), 
                each = length(epsilon)) 
  
# find the proportion of the total sample size n that the maximum
# group size takes up
  prop_n <- unlist(lapply(sizes, max))/unlist(lapply(sizes, sum))
  prop_n <- round(prop_n, digits = 2)     
  
# put together a dataframe of quantiles resulting from given parameters
  data.frame(N = rep(unlist(N_list), each = length(alphas)),
             alphas = rep(alphas, times = length(sizes)),
             type1 = type1,
             prop_n = rep(as.factor(prop_n), each = length(alphas)))
}
```

We now make use of `qq_plot` to generate figures for the paper. The first assumes equal group sizes.

```{r usage}
# make a list of sample sizes
sizes <- list(c(10, 10, 10), c(100, 100, 100), c(1000, 1000, 1000))

# simulate the plot data
kw_qqplot <- qq_plot(sizes = sizes,
         alphas = seq(from = .01, to = .99, by = .01),
         epsilon = 1,
         kw_version = "abs",
         mean = c(0, 0, 0),
         sd = c(1, 1, 1),
         reps = 1e6)

# save the plot data
save(kw_qqplot, file = "kw/data/kw_qqplot.Rda")

# create a greyscale-friendly, qualitative color palette
colors <- c("#000099", "#05e200", "#dd02e0", "#ece900")

# plot the data
kw_qqplot_plot <- ggplot(kw_qqplot) + 
    geom_line(aes(x = type1, 
                  y = alphas, 
                  col = as.factor(N)),
              size = 1.3) +
    labs(x = "Theoretical Quantiles", y = "p-value Quantiles") +
    scale_color_manual(name = "n",
                       values = colors) +
    theme_minimal(base_family = "Times",
                  base_size = 24)  + 
    theme(axis.line = element_line(colour = "black"))

kw_qqplot_plot

# save the plot
ggsave(filename = "kw_qqplot_plot.png", 
       plot = kw_qqplot_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```

The next figure uses unequal group sizes.

```{r usage}
# make a vector of sample sizes with unequal group sizes
# but identical total smaple sizes
sizes_uneq <- list(c(80, 5, 5), c(60, 15, 15), c(30, 30, 30))

# simulate the plot data
kw_qqplot_uneq <- qq_plot(sizes = sizes_uneq,
         alphas = seq(from = .01, to = .99, by = .01),
         epsilon = 1,
         kw_version = "abs",
         mean = c(0, 0, 0),
         sd = c(1, 1, 1),
         reps = 1e6)

# save the plot data
save(kw_qqplot_uneq, file = "kw/data/kw_qqplot_uneq.Rda")

# plot the data
kw_qqplot_uneq_plot <- ggplot(kw_qqplot_uneq) + 
    geom_line(aes(x = type1, 
                  y = alphas, 
                  col = as.factor(prop_n)),
              size = 1.3) +
    labs(x = "Theoretical Quantiles", y = "p-value Quantiles") +
    scale_color_manual(name = "Proportion of \nn to max(n)",
                       values = colors) +
    theme_minimal(base_family = "Times",
                  base_size = 24)  + 
    theme(axis.line = element_line(colour = "black"))

kw_qqplot_uneq_plot

# save the plot
ggsave(filename = "kw_qqplot_uneq.png", 
       plot = kw_qqplot_uneq_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```

