---
output: github_document
---
# Kruskal-Wallis Power Analysis
#### *The Effect of Ties and Tiebreaking on Power*

This file analyzes the power of our DP Kruskal-Wallis test when observations in the sample data are tied.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(plyr)
library(tidyverse)
library(rmutil)
```

Some of the functions used in this document are used, unedited, in several other documents throughout the power analysis. We load them now---documentation for each can be found in the relevant files in the `fxns` folder.

```{r}
source('kw/fxns/kw.R')
source('kw/fxns/kw_new.R')
```

Note that, as opposed to other documents in this repo, this `.Rmd` does not source the `gen_data` and `calc_stat` functions from the `fxns` folder. The framework of data generation and test statistic calculation changes slightly when we introduce ties.

`gen_data` randomly generates data to be supplied to the `kw` or `kw_new` function, as well as noise for private trials. 

* `n`, `mean`, `sd`: A vector (length h) of desired sample sizes, means, and standard deviations, respectively. Note that all three of these arguments must be of the same length.
* `mult`: A list of numerics to round values in the `x` column to, in order to introduce ties into the data.
* `reps`: The desired number of test statistics to be calculated from the output
* `epsilon`: The privacy parameter to generate the `noise` column with. Supply `epsilon = NA` for the public test (a noise column will not be generated)
* `kw_version`: Either "sq" (referring to the regular H statistic) or "abs" (referring to the unsquared absolute value H statistic) s.t. noise with the appropriate sensitivity can be generated.

```{r gen_data}
gen_data <- function(n, mult, mean, sd, reps, epsilon, kw_version) {
  if (is.na(epsilon)) { # procedure for the public test
df <-  data_frame(x = rnorm(n = reps*sum(n), 
                            mean = rep(rep(mean, n), reps),
                            sd = rep(rep(sd, n), reps)),
                  x_rounded = round_any(x, mult),
                  group = rep(rep(1:length(n), n), reps),
                  replicate = as.factor(rep(1:reps, each = sum(n))))
  } else { # procedure for the private test
      # generate noise depending on the kw_version
      if (kw_version == "sq") {
df <- data_frame(x = rnorm(n = reps*sum(n), 
                           mean = rep(rep(mean, n), reps),
                           sd = rep(rep(sd, n), reps)),
                 x_rounded = round_any(x, mult),
                 group = rep(rep(1:length(n), n), reps),
                 replicate = as.factor(rep(1:reps, each = sum(n))),
                 noise = rep(rlaplace(n = reps,
                                      m = 0,
                                      s = 78/epsilon), 
                             each = sum(n)))
      } else if (kw_version == "abs") {
df <- data_frame(x = rnorm(n = reps*sum(n), 
                       mean = rep(rep(mean, n), reps),
                       sd = rep(rep(sd, n), reps)),
                 x_rounded = round_any(x, mult),
                 group = rep(rep(1:length(n), n), reps),
                 replicate = as.factor(rep(1:reps, each = sum(n))),
                 noise = rep(rlaplace(n = reps,
                                      m = 0,
                                      s = 8/epsilon), 
                             each = sum(n)))
      }
  }
  return(list(df, kw_version))
}

```

`calc_stat` takes in a dataframe outputted by `gen_data` and simply calls the `kw` or `kw_new` function for each unique `replicate` value. The output is a distribution of test statistics.

```{r calc_stat}
calc_stat <- function(data) {
  if (data[[2]] == "abs" & "noise" %in% colnames(data[[1]])) { 
    # carry out the private abs version
    data[[1]] %>%
      dplyr::group_by(noise) %>%
      dplyr::summarize(stat = kw_new(x_rounded, group, public = FALSE) + noise[1], 
                       na.rm = TRUE) %>%
      dplyr::select(stat) %>%
      pull()
  } else if (data[[2]] == "sq" & "noise" %in% colnames(data[[1]])) { 
    # carry out the private sq version
    data[[1]] %>%
      dplyr::group_by(noise) %>%
      dplyr::summarize(stat = kw(x_rounded, group, public = FALSE) + noise[1], 
                       na.rm = TRUE) %>%
      dplyr::select(stat) %>%
      pull()    
  } else if (data[[2]] == "sq") { 
    # carry out the public sq version
    data[[1]] %>% 
      dplyr::group_by(replicate) %>%
      dplyr::summarize(stat = kw(x_rounded, group, public = TRUE)) %>%
      dplyr::select(stat) %>%
      pull()
  } else if (data[[2]] == "abs") { 
    # carry out the public abs version
    data[[1]] %>% 
      dplyr::group_by(replicate) %>%
      dplyr::summarize(stat = kw_new(x_rounded, group, public = TRUE)) %>%
      dplyr::select(stat) %>%
      pull()
  }
}
```

Because we can not explicitly specify the proportion of ties we would like to introduce into the data, we use various rounding multiples to round values in `x` to in order to introduce some proportion of ties. `prop_tied`, then, takes in the output of `gen_data` (a list of a dataframe and a `kw_version`) and finds the proportion of observations in the `x_rounded` column of the dataframe that are tied per trial of `kw`, using the metric of 1 minus the Gini coefficient of the data, and then finds the average per trial.

```{r prop_tied}
prop_tied <- function(data) {
  data[[1]] %>%
  dplyr::group_by(replicate) %>%
  dplyr::summarize(prop = 1 - sum(table(x_rounded)/length(x_rounded) * (1 - table(x_rounded)/length(x_rounded)))) %>%
  dplyr::summarize(mean(prop)) %>%
  pull()
}
```

`pwr_plot` is a wrapper function that vectorizes the functions defined above over several combinations of sample sizes and rounding multiples.

* `sizes`: A list of numeric vectors of length x, indicating the sample sizes to carry out the test with. Each element of the list must be the same length.
* `mult`: A list of numerics—the multiples to round values in `x` to in order to introduce ties into the data. 
* `epsilon`: A numeric indiciating the privacy parameter.
* `kw_version`: A character vector: either "sq" (referring to the regular H statistic), or "abs" (referring to the unsquared absolute value H statistic) so that noise with the appropriate sensitivity can be generated.
* arguments beginning with `mean` and `sd` are vectors of length x, indicating the parameters to generate data for the alternate and null distribution for.
* `reps`: A numeric indicating the number of repititions to carry out the simulations with.
* `alpha`: A numeric indicating the significance level to carry the test out with.

```{r pwr_plot}
pwr_plot <- function(sizes, mult, epsilon, kw_version, mean_null, sd_null, mean_alt, sd_alt, reps, alpha) {
   
 # we first create a list of lists, (can be thought of as a dataframe where
 # each row is a unique combination of sizes and rounding multiple)
  arg_list <- list(n =    rep(sizes, each = length(mult)), 
                   mult = rep(mult, times = length(sizes)))

# we need to do the same for the simulation of the reference distribution,
# where we assume equal group sizes
  null_sizes_1d <- lapply(sizes, mean) %>% lapply(., ceiling)
  null_sizes    <- lapply(1:length(null_sizes_1d), 
                          function(x) {rep(null_sizes_1d[[x]], 3)})  
  
  arg_list_null <- list(n =    rep(sizes, each = length(mult)), 
                        mult = rep(mult, times = length(sizes)))  
    
# simulate null distributions with varying parameters and then calculate test 
# statistics from the null distributions. the output, null_stats is a 
# dataframe, where each row is a distinct combination of sizes and rounding multiple
  
  null_stats <- pmap(.l = arg_list_null,
                 .f = gen_data,
                 mean = mean_null,
                 sd = sd_null, 
                 reps = reps,
                 epsilon = epsilon,
                 kw_version = kw_version) %>%
                map_dfc(.x = ., .f = calc_stat)

# calculate critical values for each combination of sizes and rounding multiple
  
  crit_vals  <-   map(null_stats, quantile, probs = 1 - alpha) %>% 
                  unlist(use.names = FALSE)
  
# simulate alternate distributions with varying parameters and then calculate test 
# statistics from the alternate distributions. the output, alt_stats, is a 
# dataframe, where each row is a distribution of test statistics for each
# distinct combination of sizes and rounding multiple
  
  alt_data <- pmap(.l = arg_list,
                    .f = gen_data,
                    mean = mean_alt,
                    sd = sd_alt, 
                    reps = reps,
                    epsilon = epsilon,
                    kw_version = kw_version)
  alt_stats <- map_dfc(.x = alt_data, .f = calc_stat)
  
# this is a vector of powers at given sizes and rounding multiple combinations 
   
  power_vector <- map2(.x = crit_vals, .y = alt_stats, .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)
  
# make a list of the total sample size for each element of sizes 
  N_list <- rep(lapply(sizes, sum), 
                each = length(mult))  

# find the proportion of ties in the alternate distribution
  prop_tied_vector <- map(alt_data, prop_tied) %>% unlist()
  
# put together a dataframe of powers resulting from given parameters
  plot_df <- cbind(N = unlist(N_list),
                   power_vector,
                   prop_tied_vector) %>%
             as_data_frame()

# check if there are NA's in epsilon; if so, convert to "Public" for plotting
# plot_df$epsilon[is.na(plot_df$epsilon)] <- "Public"
}
```

We now make use of the `pwr_plot` function in order to generate the plot in the paper with the varying proportion of ties.

```{r usage}
# make a vector of sample sizes
sizes_1d <- c(40, 60, 80) %>% as.list()
sizes <- lapply(1:length(sizes_1d), function(x) {rep(sizes_1d[[x]], 3)})

# simulate plot data
kw_with_ties <- pwr_plot(sizes = sizes,
         mult = list(.01, .1, .5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 5, 10),
         kw_version = "abs",
         mean_null = c(0, 0, 0),
         sd_null = c(1, 1, 1),
         mean_alt = c(-1, 0, 1),
         sd_alt = c(1, 1, 1),
         epsilon = 1,
         reps = 1e4,
         alpha = .05)

# save plot data
save(kw_with_ties, file = "kw/data/kw_with_ties.Rda")

# create a greyscale-friendly, qualitative color palette
colors <- c("#000099", "#05e200", "#dd02e0", "#ece900")

# plot the data
kw_with_ties_plot <- ggplot(kw_with_ties) + 
    geom_line(aes(x = prop_tied_vector, 
                  y = power_vector, 
                  col = as.factor(N)),
              size = 1.3) +
    labs(x = "Proportion of Ties", y = "Power") +
    scale_color_manual(name = "Total Sample \nSize n",
                       values = colors) +
    scale_linetype_discrete(name = "Test Type") +
    theme_minimal(base_family = "Times",
                  base_size = 24) +
    theme(axis.line = element_line(colour = "black"))

kw_with_ties_plot

# save the plot
ggsave(filename = "kw_with_ties.png", 
       plot = kw_with_ties_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```
