---
output: github_document
---
# Power Comparison on Real-World Data

This document outlines the procedure of the comparative power analysis of Task 2016's algorithm, our new algorithm, and the public argorithm on the NYC taxi data.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(rmutil)
```

#### The Data

This is the tidied version of the data downloaded from `http://www.andresmh.com/nyctaxitrips/`. The procedure to replicate this dataset is in `tidy_taxi.Rmd`.

```{r load_df}
load("wc/data/tidy_taxi_data.Rda")

str(tidy_taxi_data, give.attr = FALSE)
```

We want difference sets between day 1 and day 2, per driver, of: average number of passengers per trip, average time per trip (in seconds), and average distance per trip (in miles).

```{r difference_sets}
difference_sets <- tidy_taxi_data %>%
  group_by(hack_license, day) %>%
  summarize(num_passengers = mean(passenger_count),
            trip_time = mean(trip_time_in_secs),
            trip_dist = mean(trip_distance)) %>%
  ungroup() %>%
  group_by(hack_license) %>%
  summarize(num_passengers = diff(num_passengers),
            trip_time = diff(trip_time),
            trip_dist = diff(trip_dist)) %>%
  select(trip_time, trip_dist, num_passengers)

str(difference_sets)
```

#### Loading Functions

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc.R')
source('wc/fxns/wc_new.R')
source('wc/fxns/calc_task_crit_val.R')
```

#### Sample Difference Sets

We now sample, with replacement, difference sets of size n from the original data, as well as noise for each trial to add to private tests.

```{r gen_stats}
sample_sets <- function(x, n, reps, k, epsilon) {
  data_frame(x = sample(x = x, size = n*reps, replace = TRUE),
             replicate = rep(1:reps, each = n),
             noise = rep(rlaplace(n = reps,
                                  m = 0,
  # abuse indentation for 80 char limit                                                
  s = 2*n/(sqrt(n*(n+1)*(2*n+1)/6)*(epsilon))),
                         each = n),
             noise_task_priv = rep(rlaplace(n = reps,
                                            m = 0,
  # abuse indentation for 80 char limit                                                
  s = 4*k/(sqrt(2*k*(2*k+1)*(4*k+1)/6)*(epsilon))), 
                                 each = n))
}

samples <- map(difference_sets, 
               sample_sets, 
               n = 400, 
               reps = 1e5, 
               k = 15, 
               epsilon = 1)
```

#### Task 2016's Critical Values

Task 2016 details a "High Privacy" procedure to ensure a known, minimum value of `n`. In the paper, they use a `k` of 15, resulting in an `NR-min = 2*k` of 30. They do not detail how to best select a `k`, so we use `k = 15` as well.

```{r crit_val_task_high_privacy}
crit_val_task_high_privacy <- calc_task_crit_val(n = 30, 
                                                 alpha = .025, 
                                                 epsilon = 1)
```

The "High Utility"" procedure is to assume the upper bound on ties is 70%, and use the critical value from `NR_min = .3n`.

```{r crit_val_task_high_utility}
crit_val_task_high_utility <- calc_task_crit_val(n = 400*.3, 
                                                 alpha = .025, 
                                                 epsilon = 1)
```

#### Public Critical Value

We can find the public critical value with the CDF of the normal distribution.

```{r}
crit_val_public <- qnorm(p = 1 - .025)
```

#### Our Critical Value

We will use simulation to calculate a critical value for our version of the test.

The first step in simulation is to generate difference sets from the null distribution to feed to `wc`. The arguments for the function that does this, `gen_data`, are as follows:
* `n`: a numeric of desired difference set size
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations is one way of generating a null distribution of test statistics
* `sd`: a vector(length 2) of standard deviations to draw random samples with--supplying identical means and standard devations is one way of generating a null distribution of test statistics
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `k`: the priming value for which to sample the TC test's High Privacy noise from

```{r gen_data}
gen_data <- function(n, epsilon = NA, mean, sd, reps, k) {
  if (is.na(epsilon)) { # carry out the public version
    data_frame(x1 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[1], n), reps),
                          sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[2], n), reps),
                          sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n))) %>%
    select(x, replicate)
  } else { # carry out the private version
    delta_f <- n*2
    data_frame(x1 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[1], n), reps),
                          sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[2], n), reps),
                          sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n)),
               noise_task_priv = rep(rlaplace(n = reps,
                                              m = 0,
  # abuse indentation for 80 char limit                                                
  s = 4*k/(sqrt(2*k*(2*k+1)*(4*k+1)/6)*(epsilon))), 
                                     each = n),
               noise = rep(rlaplace(n = reps,
                                    m = 0,
  # abuse indentation for 80 char limit                                                
  s = 2*n/(sqrt(n*(n+1)*(2*n+1)/6)*(epsilon))), 
                                    each = n)) %>%
    select(x, replicate, noise, noise_task_priv)
  }
}
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function groups the difference sets together and supplies them to `wc_new` or `wc`, depending on the `type` argument.

```{r calc_stat}
calc_stat <- function(df, type) {
  
  reps <- n_distinct(df$replicate)
  n <- nrow(df)/reps
  denom <- sqrt(n*(n + 1)*(2*n + 1)/6)
  
    if (type == "priv") { # carry out the high privacy version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = (((wc(x))/denom) + noise_task_priv[1])) %>% 
    dplyr::select(stat) %>%
    pull()
  } else if (type == "util") { # carry out the high utility version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = (((wc(x))/denom) + noise[1])) %>% 
    dplyr::select(stat) %>%
    pull()
  } else if (type == "public") { # carry out the public version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc(x))/denom)) %>% 
    dplyr::select(stat) %>%
    pull()
  } else if (type == "new") { # carry out the new version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc_new(x))/denom) + noise[1]) %>% 
    dplyr::select(stat) %>%
    pull()
  }
}
```

Now, we make use of `gen_data` and `calc_stat` to generate a null distribution of difference sets, calculate a null distribution of test statistics, and find the critical value at a two-sided significance level `alpha = .05`.

```{r private_simulation}
# generate a null distribution of difference sets
crit_val_new <- gen_data(n = 400, 
                         epsilon = 1, 
                         mean = c(0, 0), 
                         sd = c(1, 1), 
                         reps = 1e7,
                         k = 15) %>%
  # feed these difference sets to calc_stat to generate a null 
  # distribution of test statistics
  calc_stat(., type = "new") %>%
  # find the critical value at two-sided alpha = .05
  quantile(., probs = (1 - .025), names = FALSE)
```

#### Finding Test Statistics

Make use of the `calc_stat` function to find distributions of test statistics for each of the variables for each test type.

```{r}
test_types <- c("priv", "util", "new", "public")

test_stats <- map2(rep(samples, times = length(test_types)),
                   rep(test_types, each = length(samples)),
                   calc_stat)
```

Now, we find the power of each test at two-sided `alpha = .05`, `n = 400`, for a given effect size. The input `x` to the `calc_power` function is a distribution of test statistics.

```{r calc_power}
calc_power <- function(x, crit_val) {
  mean(abs(x) > crit_val)
}

power_vector <- map2(test_stats, 
                     rep(c(crit_val_task_high_privacy,
                           crit_val_task_high_utility,
                           crit_val_new,
                           crit_val_public), each = 3),
                     calc_power) %>%
                unlist()
```

#### Plotting

Finally, we plot the data.

```{r plot}
# create a greyscale-friendly, qualitative color scheme
colors <- c("#000099", "#05e200", "#dd02e0", "#ece900")

# make a dataframe of powers
wc_vs_task_application <- data_frame(power = power_vector,
           test_type = c(rep(c("TC High \nPrivacy", 
                               "TC High \nUtility",
                               "New",
                               "Public"), 
                             each = 3)),
           effect_size = c(rep(c("Trip Time (s)", 
                                 "Trip Distance (mi)", 
                                 "Number of \nPassengers"), 
                               times = 4)))

# save the plot data
save(wc_vs_task_application, file = "wc/data/wc_vs_task_application.Rda")

# plot the dataframe
ggplot() +
  geom_bar(data = wc_vs_task_application,
           aes(x = ordered(effect_size, 
                           levels = c("Trip Time (s)", 
                                      "Trip Distance (mi)", 
                                      "Number of \nPassengers")), 
               y = power, 
               fill = ordered(test_type, 
                              levels = c("TC High \nPrivacy", 
                                         "TC High \nUtility", 
                                         "New", 
                                         "Public"),
                              labels = c("High Privacy", 
                                         "High Utility", 
                                         "New", 
                                         "Public"))),
           stat = "identity",
           position = "dodge") +
  labs(x = "Variable of Interest", 
       y = "Power") +
  theme_minimal(base_family = "Times",
                base_size = 24) +
  theme(axis.text.x = element_text(size = 16),
        legend.text = element_text(size = 14),
        axis.line = element_line(colour = "black")) +
  scale_fill_manual(name = "Test Type",
                    values = colors)
```
