---
output: github_document
---
# Comparative Power Analysis
#### *Robustness to Ties—High Utility*

This file analyzes the power of our DP Wilcoxon Signed-Rank test in comparison to Task 2016's High Utility test in the setting where sample data comes from a discrete distribution.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

This document addresses the relative robustness to ties of both our and Task 2016's differentially private version of the Wilcoxon Signed-Rank test.

```{r packages}
library(rmutil)
library(tidyverse)
```

We define a set of functions to functionalize the power analysis process, eventually coming together in a `pwr_plot` function that simulates data, calculates test statistics, finds power, and plots an informative graph based on user arguments.

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc.R')
source('wc/fxns/wc_new.R')
source('wc/fxns/gen_null.R')
source('wc/fxns/calc_task_crit_val.R')
```

`gen_data` is a function that generates difference sets and noise with given parameters. The arguments of `gen_data` are as follows:

* `n`: a numeric of desired difference set size
* `prop_0`: a numeric of the proportion of the difference set to round to 0
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations is one way of generating a null distribution of test statistics
* `sd`: a vector(length 2) of standard deviations to draw random samples with--supplying identical means and standard devations is one way of generating a null distribution of test statistics
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test

The function outputs a dataframe of distributions of difference sets.

```{r gen_data}
gen_data <- function(n, prop_0, mean, sd, reps, epsilon = NA) {
  if (is.na(epsilon)) { # carry out the public version
    vals <- data_frame(x1 = rnorm(n = round(reps*n*(1-prop_0)), 
                       mean = mean[1],
                       sd = sd[1]),
               x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                       mean = mean[2],
                       sd = sd[2]),
               x = (x1 - x2),
               replicate = as.factor(rep(1:reps, 
                                         each = round(n*(1-prop_0))))) %>%
          select(x, replicate)
    zeroes <- data_frame(x = rep(0, reps*n*prop_0),
                         replicate = rep(1:reps, each = round(n*(prop_0))))           
    rbind(vals, zeroes)
  } else { # carry out the private version
    task_s <- 2*(n*.3)/(sqrt((n*.3)*((n*.3)+1)*(2*(n*.3)+1)/6)*(epsilon)) 
    
     vals <- data_frame(x1 = rnorm(n = round(reps*(1-prop_0)*n), 
                                   mean = mean[1],
                                   sd = sd[1]),
                        x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                                  mean = mean[2],
                                  sd = sd[2]),
                        x = x1 - x2,
                        replicate = rep(1:reps, each = round(n*(1-prop_0))),
                        noise = rep(rlaplace(n = reps,
                                             m = 0,
                                             s = n*2/epsilon), 
                                    each = round(n*(1-prop_0))),
                        noise_task = rep(rlaplace(n = reps,
                                                  m = 0,
                                                  s = task_s), 
                                         each = round(n*(1-prop_0)))) %>%
          select(x, replicate, noise, noise_task)
    zeroes <- data_frame(x = rep(0, reps*n*prop_0),
                         replicate = as.factor(rep(1:reps, each = n*(prop_0))),
                         noise = rep(Inf, round(reps*n*prop_0)),
                         noise_task = rep(Inf, round(reps*n*prop_0)))
    
    rbind(vals, zeroes)
  }
}
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function simply detects whether the supplied dataframe was generated for a private or public test, and then groups the difference sets together and supplies them to `wc`. We supply a second `calc_stat` function, `calc_stat_new`, that calls `wc_new` rather than `wc`.

```{r calc_stat}
calc_stat <- function(df) {
    if ("noise_task" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc(x) - .5)/sqrt(length(x)*(length(x) + 1)*(2*length(x) + 1)/6) + noise_task[1])) %>% 
    select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc(x)) %>%
    select(stat) %>%
    pull()
  }
}

calc_stat_new <- function(df) {
    if ("noise" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x) + noise[1]) %>%
    select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x)) %>%
    select(stat) %>%
    pull()
  }
}
```

`pwr_plot` is a wrapper function for each of the functions defined above. The arguments of the function are as follows:
* `sizes`: a vector of numerics--all elements must be in the vector `c(10, 20, 30, 40, 50, 100, 200, 300, 400, 500, 1000)`
* `prop_0`: a numeric of the proportion of the difference set to round to 0
* arguments beginning with `mean` or `sd` are vectors of length 2--the first two (in position) are the arguments to generate the null distribution, and the second two are to generate the alternate distribution
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `epsilon`: NA or a numeric--the epsilon to carry the tests out with--NA indicates a public test
* `alpha`: a numeric of a significance level used in Task 2016, i.e. one of `c(.0595, .03475, .0199, .01495)`

```{r pwr_plot}
pwr_plot <- function(sizes, prop_0, mean_null, sd_null, mean_alt, sd_alt, reps, epsilon, alpha) {
   
# simulate null distributions at varying sample size and proportion 
# of zeroes combinations. null_data is a list of dataframes with 
# length equal to length(prop_0)*length(sizes)
  
  null_data <-   map2(.x = rep(sizes, each = length(prop_0)), 
                      .y = rep(prop_0, length(sizes)), 
                      .f = gen_data, 
                      mean = mean_null, 
                      sd = sd_null, 
                      reps = reps, 
                      epsilon = epsilon)
  
# simulate test statistics for the new test.
# the output, null_stats, is a dataframe, where each row is a distinct 
# combination of prop_0 & sample size, with nrow = reps

  null_stats_new <-  map_dfc(.x = rep(sizes, each = length(prop_0)), 
                             .f = gen_null,
                             epsilon = epsilon,
                             reps = reps)

# calculate critical values for each combination of proportion of 
# zeroes and sample size for the new test
  
  crit_vals_new  <-  map(null_stats_new, 
                     quantile, 
                     probs = (1 - alpha), 
                     na.rm = TRUE) %>% 
                 unlist(use.names = FALSE)
  
# find Task 2016's critical values
  
  crit_vals_task <- map(sizes*.3,
                        calc_task_crit_val,
                        alpha = alpha,
                        epsilon = 1)
  
# simulate alternate distributions at varying sample size and 
# proportion of zeroes combinations. alt_data is a list of dataframes
# with length equal to length(prop_0)*length(sizes)
  
  alt_data <-   map2(.x = rep(sizes, each = length(prop_0)), 
                     .y = rep(prop_0, length(sizes)), 
                     .f = gen_data, 
                     mean = mean_alt, 
                     sd = sd_alt, 
                     reps = reps, 
                     epsilon = epsilon)

# calculate test statistics from the alternate distributions. the output, 
# alt_stats, is a dataframe, where each row is a distinct combination 
# of prop_0 & sample size, with nrow = reps
  
  alt_stats_task <-  map_dfc(.x = alt_data, 
                        .f = calc_stat)
  
  alt_stats_new <-  map_dfc(.x = alt_data, 
                            .f = calc_stat_new)
  
# these are vectors of powers at given proportion of zeroes and 
# sample size combination for both Task's and our test
   
  power_vector_task <- map2(.x = rep(crit_vals_task, each = length(prop_0)), 
                       .y = alt_stats_task, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)
  
  power_vector_new <- map2(.x = crit_vals_new, 
                       .y = alt_stats_new, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)
  
# create a dataframe of arguments and the power_vectors and plot the data  
  plot_df <- data_frame(prop_tied = rep(prop_0, 6),
             sizes = rep(unlist(rep(sizes, each = length(prop_0))), 2),
             power = c(power_vector_task, power_vector_new),
             type = rep(c("High \nUtility", "New"), each = length(power_vector_task)))
  plot_df$power[plot_df$prop_tied == 1] <- 0

  plot_df
}
```

#### Generating the Plot

The following function call generates the plot to be used in the paper.

```{r plot}
# create a high-contrast, qualitative color scheme
colors <- c("#05e200", "#dd02e0")

# generate the plot data
wc_vs_task_hu <- pwr_plot(sizes = c(100, 300, 1000),
         prop_0 = seq(from = 0, to = 1, by = .05),
         mean_null = c(0, 0),
         sd_null = c(1, 1),
         mean_alt = c(1, 0),
         sd_alt = c(1, 1),
         reps = 1e5,
         epsilon = 1,
         alpha = .05)

# save the plot data
save(wc_vs_task_hu, file = "wc/data/wc_vs_task_hu.Rda")

# plot the data
wc_vs_task_hu_plot <- ggplot(wc_vs_task_hu) +
    aes(x = prop_tied, 
        y = power, 
        linetype = as.factor(sizes), # double map the sizes variable so that the 
        # lines are still distinguishable in black & white
        col = as.factor(type)) + 
    geom_line(size = 1.3) +
        labs(x = expression("Proportion of Zeroes in"~d[i]), y = "Power") +
        scale_color_manual(name = "Test Type",
                           values = colors,
                           labels = c("High Utility", "New")) +
        scale_linetype_manual(values = c("solid", "longdash", "dotted"),
                              name = "Database \nSize n") +
        theme_minimal(base_family = "Times", base_size = 24) +
        theme(axis.line = element_line(colour = "black"))

wc_vs_task_hu_plot

# save the plot
ggsave(filename = "wc_vs_task_hu.png", 
       plot = wc_vs_task_hu_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```