---
output: github_document
---
# Power Comparison
#### *Assuming No Ties—Comparative (incl. "Plus" Versions)*

This document outlines the procedure for the comparative power analysis of Task 2016's algorithm, our new algorithm, and the public argorithm, including the "Plus" version of Task's test, which help us to understand the comparative contributions of our innovations.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r packages}
library(tidyverse)
library(rmutil)
```

#### Loading Functions

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc.R')
source('wc/fxns/wc_new.R')
source('wc/fxns/gen_null.R')
source('wc/fxns/calc_task_crit_val.R')
```

#### Other Functions

`gen_data` is a function that generates difference sets and noise with given parameters. The arguments of `gen_data` are as follows:

* `n`: a numeric of desired difference set size
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations is one way of generating a null distribution of test statistics
* `sd`: a vector (length 2) of standard deviations to draw random samples with--supplying identical means and standard devations is one way of generating a null distribution of test statistics
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases

The function outputs a dataframe of distributions of difference sets.

```{r gen_data}
gen_data <- function(n, epsilon = NA, mean, sd, reps, k) {
  if (is.na(epsilon)) { # carry out the public version
    data_frame(x1 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[1], n), reps),
                          sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[2], n), reps),
                          sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n))) %>%
      select(x, replicate)
  } else { # carry out the private version
    delta_f <- n*2
    data_frame(x1 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[1], n), reps),
                          sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                          mean = rep(rep(mean[2], n), reps),
                          sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n)),
               noise = rep(rlaplace(n = reps,
                                    m = 0,
                                    s = delta_f/epsilon), 
                           each = n),
               noise_task_priv = rep(rlaplace(n = reps,
                                              m = 0,
  # abuse indentation for 80 char limit                                                
  s = 4*k/(sqrt(2*k*(2*k+1)*(4*k+1)/6)*(epsilon))), 
                                      each = n),
               noise_task_util = rep(rlaplace(n = reps,
                                              m = 0,
  # abuse indentation for 80 char limit                                                
  s = 2*n/(sqrt(n*(n+1)*(2*n+1)/6)*(epsilon))), 
                                     each = n)) %>%
    select(x, replicate, noise, noise_task_priv, noise_task_util)
  }
}
```

Task 2016 details a "priming" procedure to ensure a known, minimum value of `n`. `prime_data` carries out this procedure. The function takes a  dataframe outputted by `gen_data` and outputs a similar dataframe.

```{r prime_data}
prime_data <- function(df, k) {
      reps <- n_distinct(df$replicate)
      n <- nrow(df)/reps
      prime <- data_frame(x = rep(c(rep(max(abs(df$x) + 1), k), 
                                    rep(-max(abs(df$x) - 1), k)), 
                                  reps),
                          replicate = rep(1:reps, each = 2*k),
                          noise = rep(Inf, 2*k*reps),
                          noise_task_priv = rep(Inf, 2*k*reps),
                          noise_task_util = rep(Inf, 2*k*reps))
      rbind(df, prime)
} 
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function simply detects whether the supplied dataframe was generated for a private or public test, and then groups the difference sets together and supplies them to `wc_new`.

```{r calc_stat}
calc_stat <- function(df, type) {
    if (type == "priv") { # carry out the high privacy version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc(x))/sqrt(length(x)*(length(x) + 1)*(2*length(x) + 1)/6) + noise_task_priv[1])) %>% 
    dplyr::select(stat) %>%
    pull()
  } else if (type == "util") { # carry out the high utility version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc(x))/sqrt(length(x)*(length(x) + 1)*(2*length(x) + 1)/6) + noise_task_util[1])) %>% 
    dplyr::select(stat) %>%
    pull()
  } else if (type == "public") { # carry out the public version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(., stat = ((wc(x))/sqrt(length(x)*(length(x) + 1)*(2*length(x) + 1)/6))) %>% 
    dplyr::select(stat) %>%
    pull()
  }
}

calc_stat_new <- function(df) {
    if ("noise" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x) + noise[1]) %>%
    dplyr::select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x)) %>%
    dplyr::select(stat) %>%
    pull()
  }
}
```

`pwr_plot` is a wrapper function for the functions defined above. The arguments of the function are as follows:
* `sizes`: a list of numerics
* `epsilon`: a numeric (currently only supports `epsilon = 1`)
* arguments beginning with `mean` or `sd` are vectors of length 2--the first two (in position) are the arguments to generate the null distribution, and the second two are to generate the alternate distribution
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `alpha`: a numeric, the two-sided significance level
* `k`: the priming value to use for Task 2016 High Privacy variant

```{r pwr_plot}
pwr_plot <- function(sizes, epsilon, mean_null, sd_null, mean_alt, sd_alt, reps, alpha, k) {
 
# Find Power of the New Test -------------------------------------------    
      
# generate a null distribution of test statistics
  
  null_stats_new <- map_dfc(sizes,
                    gen_null,
                    epsilon = epsilon,
                    reps = reps) %>%
                    abs()

# calculate critical values for the new test for each sample size
  
  crit_vals_new  <-   map(null_stats_new, 
                          quantile, 
                          probs = (1 - alpha)) %>% 
                      unlist(use.names = FALSE)
  
# simulate alternate distributions at varying sample sizes, and then
# calculate test statistics from the null distributions. the output, 
# alt_stats_new, is a dataframe, where each row is a distribution of 
# test statistics for a given sample size, with nrow = reps
  
  alt_data <-     map(.x = sizes, 
                      .f = gen_data, 
                      epsilon = epsilon,
                      mean = mean_alt, 
                      sd = sd_alt, 
                      reps = reps,
                      k = 15)
  
  alt_stats_new <- map_dfc(.x = alt_data, 
                           .f = calc_stat_new) %>%
                  abs()
  
# this is a vector of powers at given epsilon and sample size combinations 
   
  power_vector_new <- map2(.x = crit_vals_new, 
                       .y = alt_stats_new, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)
  
# Find Power of the Public Test ----------------------------------------
  
# simulate null distributions at varying sample sizes, and then
# calculate test statistics from the null distributions. the output, 
# null_stats_public, is a dataframe, where each row is a distribution 
# at a given sample size, with nrow = reps
  
  null_data_public <- map(.x = sizes, 
                          .f = gen_data, 
                          epsilon = NA, 
                          mean = mean_null, 
                          sd = sd_null, 
                          reps = reps,
                          k = k)
  
  null_stats_public <- map_dfc(.x = null_data_public, 
                               .f = calc_stat,
                               type = "public") %>%
                      abs()

# calculate critical values for the public test for each sample size
  
  crit_vals_public  <-   map(null_stats_public, 
                             quantile, 
                             probs = (1 - alpha)) %>% 
                         unlist(use.names = FALSE)
  
# simulate public alternate distributions at varying sample sizes 
  alt_data_public  <- map(.x = sizes, 
                          .f = gen_data, 
                          epsilon = NA,
                          mean = mean_alt, 
                          sd = sd_alt, 
                          reps = reps,
                          k = k)
  
# calculate test statistics from the null distributions. the output, 
# alt_stats_public, is a dataframe, where each row is a distribution 
# of test statistics for a given sample size, with nrow = reps
  
  alt_stats_public <- map_dfc(.x = alt_data_public, 
                              .f = calc_stat,
                              type = "public") %>%
                      abs()
  
# this is a vector of powers at given epsilon and sample size combinations 
   
  power_vector_public <- map2(.x = crit_vals_public, 
                       .y = alt_stats_public, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)

# Find the Power of the High Privacy and High Utility algorithms ------------------
   
# prime alt_data with 2*k values  
  alt_data_primed <- map(alt_data, prime_data, k = k)
  
# generate an alternate distribution of test statistics for the high privacy test
  alt_stats_task_priv <- map_dfc(.x = alt_data_primed, 
                           .f = calc_stat,
                           type = "priv") %>%
                        abs()
  
# calculate crit vals of tasks high privacy test
  crit_vals_task_priv <- calc_task_crit_val(n = 2*k,
                                       alpha = alpha,
                                       epsilon = epsilon) %>%
                    rep(., length(sizes))
 
# find power   
  power_vector_task_priv <- map2(.x = crit_vals_task_priv, 
                       .y = alt_stats_task_priv, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE) 

# generate an alternate distribution of test statistics for the high utility test
  alt_stats_util <- map_dfc(.x = alt_data, 
                           .f = calc_stat,
                           type = "util")  %>%
                  abs()
  
# calculate crit vals of tasks high utility test 
 crit_vals_task_util <- map(unlist(sizes)*.3,
                            calc_task_crit_val,
                            alpha = alpha,
                            epsilon = 1) %>%
                        unlist()

# find power
 power_vector_task_util <- map2(.x = crit_vals_task_util, 
                                .y = alt_stats_util, 
                                .f = function(x, y) {mean(y > x)}) %>% 
                                unlist(use.names = FALSE)  

# Find The Task "High Utility Plus" Power -------------------------
 
 # simulate null distributions at varying sample sizes
  
  null_data_util <-map(.x = round(unlist(sizes)*.3), 
                       .f = gen_data, 
                       epsilon = epsilon, 
                       mean = mean_null, 
                       sd = sd_null, 
                       reps = reps,
                       k = k)
 
# calculate test statistics for the Task High Utility Plus test from 
# the null distributions    
 
  null_stats_util <- map_dfc(.x = null_data_util, 
                            .f = calc_stat,
                            type = "util") %>%
                    abs()

# calculate critical values for the High Utility Plus test for each n
  
  crit_vals_new_util<-map(null_stats_util, 
                          quantile, 
                          probs = (1 - alpha)) %>% 
                      unlist(use.names = FALSE) 

# find power
 power_vector_task_util_plus <- map2(.x = crit_vals_new_util, 
                                .y = alt_stats_util, 
                                .f = function(x, y) {mean(y > x)}) %>% 
                                unlist(use.names = FALSE)   
 
# Find The Task "High Privacy Plus" Power -------------------------

 # simulate null distributions at varying sample sizes
  null_data <-     map(.x = sizes, 
                       .f = gen_data, 
                       epsilon = epsilon, 
                       mean = mean_null, 
                       sd = sd_null, 
                       reps = reps,
                       k = k)
   
# prime null_data with 2*k values  
  null_data_primed <- map(null_data, prime_data, k = k)
  
# generate an null distribution of test statistics for the high privacy test
  null_stats_task_priv <- map_dfc(.x = null_data_primed, 
                           .f = calc_stat,
                           type = "priv") %>%
                          abs()
 
# calculate critical values for each sample size
  
  crit_vals_new_priv <-  map(null_stats_task_priv, 
                             quantile, 
                             probs = (1 - alpha)) %>% 
                         unlist(use.names = FALSE)  
  
# find power
 power_vector_task_priv_plus <- map2(.x = crit_vals_new_priv, 
                                .y = alt_stats_task_priv, 
                                .f = function(x, y) {mean(y > x)}) %>% 
                                unlist(use.names = FALSE)    
  
# Combine the Data and Plot it -------------------------------------
    
# create a dataframe of arguments and resultant powers, plot it  
  plot_df <- data.frame(power = c(power_vector_task_priv,
                                  power_vector_task_util,
                                  power_vector_task_priv_plus,
                                  power_vector_task_util_plus,
                                  power_vector_new,
                                  power_vector_public),
                        sizes = rep(unlist(sizes), 6),
                        type = rep(c("High Privacy",
                                     "High Utility",
                                     "High Privacy +",
                                     "High Utility +",
                                     "New",
                                     "Public"), 
                                   each = length(sizes)))
}
```

The following function call generates the plot to be used in the paper.

```{r plot}
sizes <- c(5, seq(from = 10, to = 200, by = 10)) %>% as.list()

# generate the plot data
wc_comp_crit_val_w_plus <- pwr_plot(sizes = sizes, 
         epsilon = 1,
         mean_null = c(0, 0), 
         sd_null = c(1, 1), 
         mean_alt = c(1, 0), 
         sd_alt = c(1, 1),
         reps = 1e6, 
         alpha = .05,
         k = 15)

# save the plot data
save(wc_vs_task_w_plus, file = "wc/data/wc_vs_task_w_plus.Rda")

# create a high-contrast, grey-scale friendly, qualitative color scheme
colors <- c("#000686", "#05e200", "#cc6600", "#0f4f50", "#dd02e0", "#ece900")

# plot the data
wc_vs_task_w_plus_plot <- ggplot(wc_vs_task_w_plus) +
             geom_line(aes(x = sizes, 
                           y = power, 
                           linetype = ordered(type,
                                                levels = c("High Privacy",
                                                           "High Utility",
                                                           "High Privacy +",
                                                           "High Utility +",
                                                           "New",
                                                           "Public")), 
                           col = ordered(type,
                                           levels = c("High Privacy",
                                                      "High Utility",
                                                      "High Privacy +",
                                                      "High Utility +",
                                                      "New",
                                                      "Public"))),
                       size = 1.3) +
             labs(x = expression("Database Size"~italic("n")), 
                  y = "Power") +
             #scale_x_log10() + 
             theme_minimal(base_family = "Times", base_size = 24) +
             scale_linetype_manual(name = "Test Type",
                                   values = c(5, 2, 6, 6, 3, 1)) +
             scale_color_manual(name = "Test Type", 
                                values = colors) +
             theme(axis.line = element_line(colour = "black"))

wc_vs_task_w_plus_plot 

# save the plot
ggsave(filename = "wc_vs_task_w_plus.png", 
       plot = wc_vs_task_w_plus_plot ,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```

