---
output: github_document
---
# Noncomparative Power Analysis
#### *Multivariate Analysis Assuming No Ties*

This file analyzes the power of our DP Wilcoxon Signed-Rank test varying sample size and epsilon, where sample data is assumed to come from a continuous distribution.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r packages}
library(tidyverse)
library(rmutil)
library(latex2exp)
```

We define a set of functions to functionalize the power analysis process, eventually coming together in a `pwr_plot` function that simulates data, calculates test statistics, finds power, and plots an informative graph based on user arguments.

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc_new.R')
source('wc/fxns/gen_null.R')
```

`gen_data` is a function that generates difference sets and noise with given parameters. The arguments of `gen_data` are as follows:

* `n`: a numeric of desired difference set size
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations generates a null distribution of difference sets
* `sd`: a vector(length 2) of standard deviations to draw random samples with--supplying identical means and standard devations generates a null distribution of difference sets
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases

The function outputs a dataframe of distributions of difference sets.

```{r gen_data}
gen_data <- function(n, epsilon = NA, mean, sd, reps) {
  if (is.na(epsilon)) { # carry out the public version
    data_frame(x1 = rnorm(n = reps*n, 
                       mean = rep(rep(mean[1], n), reps),
                       sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                       mean = rep(rep(mean[2], n), reps),
                       sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n))) %>%
      select(x, replicate)
  } else { # carry out the private version
    delta_f <- n*2
    data_frame(x1 = rnorm(n = reps*n, 
                       mean = rep(rep(mean[1], n), reps),
                       sd = rep(rep(sd[1], n), reps)),
               x2 = rnorm(n = reps*n, 
                       mean = rep(rep(mean[2], n), reps),
                       sd = rep(rep(sd[2], n), reps)),
               x = x1 - x2,
               replicate = as.factor(rep(1:reps, each = n)),
               noise = rep(rlaplace(n = reps,
                                    m = 0,
                                    s = delta_f/epsilon), 
                       each = n)) %>%
    select(x, replicate, noise)
  }
}
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function simply detects whether the supplied dataframe was generated for a private or public test, and then groups the difference sets together and supplies them to `wc`.

```{r calc_stat}
calc_stat <- function(df) {
    if ("noise" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc(x) + noise[1]) %>%
    dplyr::select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc(x)) %>%
    dplyr::select(stat) %>%
    pull()
  }
}
```

`pwr_plot` is a wrapper function for each of the functions defined above. The arguments of the function are as follows:
* `sizes`: a list of numerics
* `epsilons`: a list of numerics (can include NA)
* arguments beginning with `mean` or `sd` are vectors of length 2--the first two (in position) are the arguments to generate the null distribution, and the second two are to generate the alternate distribution
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases

```{r pwr_plot}
pwr_plot <- function(sizes, epsilons, mean_null, sd_null, mean_alt, sd_alt, reps) {
   
# generate a distribution of null test statistics at each sample size and
# epsilon combination
  
  null_stats <-   map2_dfc(.x = rep(sizes, each = length(epsilons)), 
                           .y = rep(epsilons, times = length(sizes)), 
                           .f = gen_null, 
                           reps = reps) %>%
                  abs()

# calculate critical values for each combination of epsilon and sample size
  
  crit_vals  <-   map(null_stats, 
                      quantile, 
                      probs = (1 -.05)) %>% 
                  unlist(use.names = FALSE)
  
# simulate alternate distributions of difference sets at varying sample 
# size and epsilon combinations, and then calculate test statistics 
# from the distributions. the output, alt_stats, is a dataframe, where each 
# row is a distinct combination of epsilons & sample size, with nrow = reps
  
  alt_stats <-    map2(.x = rep(sizes, each = length(epsilons)), 
                       .y = rep(epsilons, times = length(sizes)), 
                       .f = gen_data, 
                       mean = mean_alt, sd = sd_alt, reps = reps) %>%
                  map_dfc(.x = ., .f = calc_stat) %>%
                  abs()
  
# this is a vector of powers at given epsilon and sample size combinations 
   
  power_vector <- map2(.x = crit_vals, 
                       .y = alt_stats, 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)

# check if there are NA's in epsilons; if so, convert to "Public" for plotting
  epsilons[is.na(epsilons)] <- "Public"
    
# create a dataframe of arguments and resultant powers, plot it  
  plot_df <- data_frame(epsilon = unlist(rep(epsilons, length(sizes))),
             sizes = unlist(rep(sizes, each = length(epsilons))),
             power = power_vector)

  plot_df
}
```

The following function calls generate the plot to be used in the paper.

```{r plot}
# these sample sizes are chosen so as to create equal intervals between 
# points in the x-axis when plotting log-10 sample size
sizes <- seq(from = 1.45, to = 2.45, by = (.05))^10 %>% round_any(., 30) %>% c(10, .) %>% as.list()
epsilons <- list(.01, .1, 1, NA)

# create a greyscale-friendly, qualitative color scheme
colors <- c("#05e200", "#000099", "#dd02e0", "#ece900")

# generate plot data
wc_no_ties <- pwr_plot(sizes = sizes, 
         epsilons = epsilons,
         mean_null = c(0, 0), 
         sd_null = c(1, 1), 
         mean_alt = c(0, 1), 
         sd_alt = c(1, 1), 
         reps = 1e6)

# save the plot data
save(wc_no_ties, file = "wc/data/wc_no_ties.Rda")

# plot the data
wc_no_ties_plot <- ggplot(wc_no_ties) +
             geom_line(aes(x = sizes, 
                           y = power, 
                           linetype = as.factor(epsilon), 
                           col = as.factor(epsilon)),
                       size = 1.3) +
             labs(x = expression("Database Size"~italic("n")), 
                  y = "Power") +
             scale_x_log10() + 
             theme_minimal(base_family = "Times", base_size = 24) +
             scale_linetype_manual(values = c(4, 2, 3, 1),
                                   name = TeX("$\\epsilon$")) +
             scale_color_manual(name = TeX("$\\epsilon$"), 
                                values = colors) +
             theme(axis.line = element_line(colour = "black"))

# save the plot
ggsave(filename = "wc_no_ties.png", 
       plot = wc_no_ties_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```