---
output: github_document
---
# Noncomparative Power Analysis
#### Generating the *qqplot*s

This file analyzes the error rates of our DP Wilcoxon Signed-Rank test using quantile-quantile plots.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r packages}
library(plyr)
library(rmutil)
library(tidyverse)
library(latex2exp)
```

We define a set of functions to functionalize the power analysis process, eventually coming together in a `pwr_plot` function that simulates data, calculates test statistics, finds power, and plots an informative graph based on user arguments.

#### Loading Functions

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc_new.R')
source('wc/fxns/gen_null.R')
source('wc/fxns/calc_task_crit_val.R')
```

`gen_data` is a function that generates difference sets and noise with given parameters. The arguments of `gen_data` are as follows:

* `n`: a numeric of desired difference set size
* `prop_0`: a numeric of the proportion of the difference set to round to 0
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations generates a null distribution of test statistics
* `sd`: a vector(length 2) of standard deviations to draw random samples with--supplying identical means and standard devations generates a null distribution of test statistics
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test

The function outputs a dataframe of distributions of difference sets.

```{r gen_data}
gen_data <- function(n, prop_0, mean, sd, reps, epsilon) {
  
  if (is.na(epsilon)) { # carry out the public version
    
    vals <- data_frame(x1 = rnorm(n = round(reps*n*(1-prop_0)), 
                       mean = mean[1],
                       sd = sd[1]),
               x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                       mean = mean[2],
                       sd = sd[2]),
               x = (x1 - x2),
               replicate = as.factor(rep(1:reps, 
                                         each = round(n*(1-prop_0))))) %>%
          select(x, replicate)
    
    zeroes <- data_frame(x = rep(0, 
                                 times = round(reps*n*prop_0)),
                         replicate = rep(1:reps, 
                                         each = round(n*(prop_0))))           
    
    rbind(vals, zeroes)
    
  } else { # carry out the private version
    
    task_s <- 2*(n*.3)/(sqrt((n*.3)*((n*.3)+1)*(2*(n*.3)+1)/6)*(epsilon)) 
    
    vals <- data_frame(x1 = rnorm(n = round(reps*(1-prop_0)*n), 
                                  mean = mean[1],
                                  sd = sd[1]),
                       x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                                 mean = mean[2],
                                 sd = sd[2]),
                       x = x1 - x2,
                       replicate = rep(1:reps, 
                                       each = round(n*(1-prop_0))),
                       noise = rep(rlaplace(n = reps,
                                            m = 0,
                                            s = n*2/epsilon), 
                                    each = round(n*(1-prop_0)))) %>%
          select(x, replicate, noise)
     
    zeroes <- data_frame(x = rep(0, round(reps*n*prop_0)),
                         replicate = as.factor(rep(1:reps, 
                                                   each = round(n*(prop_0)))),
                         noise = rep(Inf, 
                                     times = round(reps*n*prop_0)))
    
    rbind(vals, zeroes)
  }
}
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function simply detects whether the supplied dataframe was generated for a private or public test, and then groups the difference sets together and supplies them to `wc_new`.

```{r calc_stat}
calc_stat <- function(df) {
    if ("noise" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x) + noise[1]) %>%
    dplyr::select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x)) %>% 
    dplyr::select(stat) %>%
    pull()
  }
}
```

`pwr_plot` is a wrapper function for each of the functions defined above. The arguments of the function are as follows:

* `alphas`: a vector of numerics indicating the significance levels to examine the test over. `seq(from = .01, to = .99, by = .01)` is a sensible default, but to make the procedure less computationally intensive, one can increase the `by` argument. The test assumes that all of these are two-sided alphas.
* `sizes`: a list of numerics-the sample sizes *n* to use for the tests
* `epsilon`: a numeric
* `prop_0`: the proportion of zeroes to add to the difference set of one of the lines
* `prop_0_2`: the proportion of zeroes to add to the difference set of another line
* arguments beginning with `mean` or `sd` are vectors of length 2--the arguments to generate the distributions (for qqplots, the values in the vectors should be identical so as to generate a null distribution)
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases

```{r pwr_plot}
pwr_plot <- function(alphas, size, epsilon, prop_0, prop_0_2, mean, sd, reps) {
 
# Finding null critical values ---------------------------------  
  null_stats <- gen_null(size, epsilon, reps) %>% abs()
  
  crit_vals <- quantile(null_stats, 
                        probs = 1 - alphas, 
                        na.rm = TRUE, 
                        names = FALSE)

# Find Power with  prop_0 ---------------------------------------- 
  alt_data_w_ties <- gen_data(size, epsilon, mean = mean, 
                              sd = sd, reps = reps, prop_0 = prop_0)
  
  alt_stats_w_ties <- calc_stat(alt_data_w_ties) %>% abs()
  
  t1_w_ties <- map(.x = crit_vals,
            .f = function(x, y) {mean(y > x)},
            y = alt_stats_w_ties) %>%
    unlist(use.names = FALSE)

# Find power with prop_0_2 ---------------------------------------
  alt_data_w_more_ties <- gen_data(size, epsilon, mean = mean, 
                                   sd = sd, reps = reps, prop_0 = prop_0_2)
  
  alt_stats_w_more_ties <- calc_stat(alt_data_w_more_ties) %>% abs()
  
  t1_w_more_ties <- map(.x = crit_vals,
            .f = function(x, y) {mean(y > x)},
            y = alt_stats_w_more_ties) %>%
    unlist(use.names = FALSE)

# Find power with no zeroes ---------------------------------------   
  alt_data_no_ties <- gen_data(size, epsilon, mean = mean, 
                               sd = sd, reps = reps, prop_0 = 0)
  
  alt_stats_no_ties <- calc_stat(alt_data_no_ties) %>% abs()
  
  t1_no_ties <- map(.x = crit_vals,
            .f = function(x, y) {mean(y > x)},
            y = alt_stats_no_ties) %>%
    unlist(use.names = FALSE)
  
  plot_df <- data.frame(alphas = rep(alphas, 3),
                        quant_p = c(t1_w_ties, t1_w_more_ties, t1_no_ties),
                        type = rep(c("w_ties", "w_more_ties", "no_ties"), each = length(alphas)))

  plot_df
}
```

#### Generating the Plot

The following function calls generate the plot to be used in the paper.

```{r plot}
# create a high-contrast, qualitative color scheme
colors <- c("#05e200", "#dd02e0", "#000099", "#ece900")

# generate the plot data
wc_qqplot <- pwr_plot(alphas = seq(from = .01, to = .99, by = .01),
         size = 500,
         epsilon = 1,
         prop_0 = .3,
         prop_0_2 = .9,
         mean = c(0, 0),
         sd = c(1, 1),
         reps = 1e6)

# save the plot data
save(wc_qqplot, file = "wc/data/wc_qqplot.Rda")


plot_labels <- c(expression('0% Zeroes in d'[i]),
                    expression('30% Zeroes in d'[i]),
                    expression('90% Zeroes in d'[i]))

# plot the data  
wc_qqplot_plot <- ggplot(wc_qqplot) + 
    geom_line(aes(x = quant_p, 
                  y = alphas, 
                  linetype = factor(type, 
                                    levels = c("no_ties", "w_ties", "w_more_ties"),
                                    labels = plot_labels),
                  col = factor(type, 
                                    levels = c("no_ties", "w_ties", "w_more_ties"),
                                    labels = plot_labels)),
              size = 1.3) +
        lims(x = c(0, 1), y = c(0, 1)) +
        labs(x = "Theoretical Quantiles", y = "p-value Quantiles") +
        scale_color_manual(name = "Test Type",
                           values = colors,
                           labels = plot_labels) +
        scale_linetype_manual(name = "Test Type",
                                values = c(1, 2, 3, 4),
                           labels = plot_labels) +
        theme_minimal(base_family = "Times", base_size = 24) +
        theme(axis.line = element_line(colour = "black"))

# save the plot
ggsave(filename = "wc_qqplot.png", 
       plot = wc_qqplot_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```
