---
output: github_document
---
# NYC Taxi Data Tidying

This file serves to demonstrate the procedure for tidying and filtering the sample dataset used in our own and Task 2016's paper. The result, `tidy_taxi_data.Rda`, is a tidy dataset containing relevant variables on all taxi trips on January 1st and 2nd in 2013 in New York City.

#### Setup

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
```

The original data file is available at \link{http://www.andresmh.com/nyctaxitrips/}. The relevant file is `trip_data_1.csv`.

```{r import}
# this line will need to be changed in respect to where the trip_data_1.csv file is stored
raw_trip_data <- read_csv("trip_data_1.csv")

str(raw_trip_data, give.attr = FALSE)
```

Isolate rides that both occurred in the first two days of January and were given by drivers who gave rides on both the 1st and the 2nd.

```{r filter}
tidy_taxi_data <- raw_trip_data %>%
  filter(day(pickup_datetime) <= 2 & day(dropoff_datetime) <= 2) %>%
  mutate(dom = day(pickup_datetime)) %>%
  filter(dom <= 2) %>%
  group_by(hack_license) %>%
  mutate(ndays = n_distinct(dom)) %>%
  filter(ndays == 2)
  
n_distinct(tidy_taxi_data$hack_license)
```

Task 2016 claims there are 17,069 distinct drivers in the dataset meeting these criteria--we found there to be 17,066.

The variables of interest are `hack_license`, `day`, `passenger_count`, `trip_time_in_secs`, and `trip_distance`. We select only those variables.

```{r select}
tidy_taxi_data <- tidy_taxi_data %>%
  select(hack_license, 
         day = dom, 
         passenger_count, 
         trip_time_in_secs, 
         trip_distance)
```

Finally, we save the file to be used in power analysis.

```{r save}
save(tidy_taxi_data, 
     file = "wc/data/tidy_taxi_data.Rda")
```

