---
output: github_document
---
# Noncomparative Power Analysis
#### *The Effect of Ties and Tiebreaking on Power*

This file analyzes the power of our DP Wilcoxon Signed-Rank test in the setting where sample data comes from a discrete distribution.

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r packages}
library(plyr)
library(rmutil)
library(tidyverse)
library(latex2exp)
```

We define a set of functions to functionalize the power analysis process, eventually coming together in a `pwr_plot` function that simulates data, calculates test statistics, finds power, and plots an informative graph based on user arguments.

#### Loading Functions

Some functions used throughout this repo are universal to most of the files. We load them now. Refer to the relevant files in the `fxns` folder for documentation.

```{r fxns}
source('wc/fxns/wc_new.R')
source('wc/fxns/gen_null.R')
```

`gen_data` is a function that generates difference sets and noise with given parameters. Note that this is a different `gen_data` than used throughout the rest of the `wc` repo. The arguments of `gen_data` are as follows:

* `n`: a numeric of desired difference set size
* `prop_0`: a numeric of the proportion of the difference set to round to 0
* `mean`: a vector (length 2) of means to draw random samples with--supplying identical means and standard deviations generates a null distribution of test statistics
* `sd`: a vector (length 2) of standard deviations to draw random samples with--supplying identical means and standard devations generates a null distribution of test statistics
* `reps`: a numeric indicating the number of times to repeat the simulation--as reps increases, precision increases
* `epsilon`: NA or a numeric--the epsilon to carry the test out with--NA indicates a public test

The function outputs a dataframe of distributions of difference sets.

```{r gen_data}
gen_data <- function(n, prop_0, mean, sd, reps, epsilon) {
  
  if (is.na(epsilon)) { # carry out the public version
    
    vals <- data_frame(x1 = rnorm(n = round(reps*n*(1-prop_0)), 
                                  mean = mean[1],
                                  sd = sd[1]),
               x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                          mean = mean[2],
                          sd = sd[2]),
               x = (x1 - x2),
               replicate = as.factor(rep(1:reps, 
                                         each = round(n*(1-prop_0))))) %>%
          select(x, replicate)
    
    zeroes <- data_frame(x = rep(0, 
                                 times = round(reps*n*prop_0)),
                         replicate = rep(1:reps, 
                                         each = round(n*(prop_0))))           
    
    rbind(vals, zeroes)
    
  } else { # carry out the private version
    
    task_s <- 2*(n*.3)/(sqrt((n*.3)*((n*.3)+1)*(2*(n*.3)+1)/6)*(epsilon)) 
    
    vals <- data_frame(x1 = rnorm(n = round(reps*(1-prop_0)*n), 
                                  mean = mean[1],
                                  sd = sd[1]),
                       x2 = rnorm(n = round(reps*n*(1-prop_0)), 
                                 mean = mean[2],
                                 sd = sd[2]),
                       x = x1 - x2,
                       replicate = rep(1:reps, 
                                       each = round(n*(1-prop_0))),
                       noise = rep(rlaplace(n = reps,
                                            m = 0,
                                            s = n*2/epsilon), 
                                    each = round(n*(1-prop_0)))) %>%
          select(x, replicate, noise)
     
    zeroes <- data_frame(x = rep(0, round(reps*n*prop_0)),
                         replicate = as.factor(rep(1:reps, 
                                                   each = round(n*(prop_0)))),
                         noise = rep(Inf, 
                                     times = round(reps*n*prop_0)))
    
    rbind(vals, zeroes)
  }
}
```

`calc_stat` is a function that takes in a dataframe outputted by `gen_data` and outputs a vector of test statistics with length `reps`. The function simply detects whether the supplied dataframe was generated for a private or public test, and then groups the difference sets together and supplies them to `wc_new`.

```{r calc_stat}
calc_stat <- function(df) {
    if ("noise" %in% colnames(df)) { # carry out the private version
  df %>%
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x) + noise[1]) %>%
    dplyr::select(stat) %>%
    pull()
  } else { # carry out the public version
  df %>% 
    dplyr::group_by(replicate) %>%
    dplyr::summarize(stat = wc_new(x)) %>% 
    dplyr::select(stat) %>%
    pull()
  }
}
```

The `pwr_plot` function is a wrapper for each of the functions defined above. It maps the functions over user arguments and returns a dataframe that can be plotted to make a power plot. The arguments are as follows:

* `sizes`: a list of numerics indicating the sample sizes to run the test on
* `epsilons`: a list of numerics indicating the privacy parameters to the run the test on, where NA indicates the public test
* `prop_0`: the proportion of zeroes to introduce into the data
* arguments beginning with `mean` and `sd` are vectors of length 2 indicating the parameters to generate the *u* and *v* sets with. 
* `reps`: the number of simulations to carry the test out with

```{r pwr_plot}
pwr_plot <- function(sizes, epsilons, prop_0, mean_null, sd_null, mean_alt, sd_alt, reps) {
 
# generate a distribution of null test statistics at each sample size and
# epsilon combination 
  null_stats <- map2_dfc(.x = rep(unlist(sizes), times = length(epsilons)),
                    .y = rep(unlist(epsilons), each = length(sizes)),
                    .f = gen_null,
                    reps = reps)
  
# find two-sided critical values
  crit_vals  <-  map(.x = abs(null_stats), 
                     .f = quantile, 
                     probs = (1 - .05), 
                     na.rm = TRUE) %>% 
                 unlist(use.names = FALSE)
  

# simulate alternate distributions of difference sets at varying sample 
# size and epsilon combinations, and then calculate test statistics 
# from the distributions. the output, alt_stats, is a dataframe, where each 
# row is a distinct combination of epsilons & sample size, with nrow = reps
    
  alt_data <-   map2(.x = rep(unlist(sizes), times = length(epsilons)), 
                     .y = rep(unlist(epsilons), each = length(sizes)), 
                     .f = gen_data, 
                     mean = mean_alt, 
                     sd = sd_alt, 
                     reps = reps, 
                     prop_0 = prop_0)

  alt_stats <-  map_dfc(.x = alt_data, 
                        .f = calc_stat)
  
# calculate power of the tests 
  power_vector <- map2(.x = crit_vals, 
                       .y = abs(alt_stats), 
                       .f = function(x, y) {mean(y > x)}) %>% 
                       unlist(use.names = FALSE)
  
# convert NA epsilons to "Public" for plotting  
  epsilons[is.na(epsilons)] <- "Public"
  
# make a dataframe of arguments and resultant powers  
  plot_df <- data.frame(sizes = rep(sizes, times = length(epsilons)),
                        epsilons = rep(unlist(epsilons), each = length(sizes)),
                        power = power_vector)
  
  plot_df
}
```

#### Generating the Plot

The following function call generates the plot to be used in the paper.

```{r plot}
# create a greyscale-friendly, qualitative color scheme
colors <- c("#05e200", "#000099", "#dd02e0", "#ece900")

# generate the plot data
wc_30pct_ties <- pwr_plot(sizes = seq(from = 1.45, 
                                       to = 2.45, 
                                       by = (.05))^10 %>% round_any(., 30) %>% c(10, .),
         epsilons = list(.01, .1, 1, NA),
         prop_0 = .3,
         mean_null = c(0, 0),
         sd_null = c(1, 1),
         mean_alt = c(0, 1),
         sd_alt = c(1, 1),
         reps = 1e6)

# save the plot data
save(wc_30pct_ties, file = "wc/data/wc_30pct_ties.Rda")

# plot the data  
wc_30_pct_ties_plot <- ggplot(wc_30pct_ties) + 
        geom_line(aes(x = sizes, 
                      y = power, 
                      linetype = as.factor(epsilons),
                      col = as.factor(epsilons)),
                  size = 1.3) +
        labs(x = expression("Database Size"~italic("n")), y = "Power") +
        scale_color_manual(name = TeX('$\\epsilon$'),
                           values = colors) +
        scale_x_log10() +
        scale_linetype_manual(name = TeX('$\\epsilon$'),
                              values = c(4, 2, 3, 1)) +
        theme_minimal(base_family = "Times", base_size = 24) +
        theme(axis.line = element_line(colour = "black"))

# save the plot
ggsave(filename = "wc_30_pct_ties.png", 
       plot = wc_30_pct_ties_plot,
       path = "figures",
       width = 10,
       height = 6,
       units = "in")
```
